Towards an ecology of vagueness

### 2017-11-11 book_chapter/paper.tex
Addressing Michael's comments.

In the following, we lay down the formal details of our model.
Let's start by defining $A$ to be the set of imprecision values considered.
For each value $\alpha \in A$, the proportion of its population is given by $P(\alpha)$, such that $\sum_{\alpha \in A} P(\alpha) = 1$.
The probability that a given agent with imprecision $\alpha$ (or of type $\alpha$) observes $t_o$ when the actual state is $t_a$ is given by $P_{o}^{\alpha}(t_{o}|t_{a})$.
If the same agent intends to realize $t_i$, the probability that he actually realizes $t_r$ instead is given by $P_{r}^{\alpha}(t_{r}|t_{i})$.
Following~\textcite[26]{franke_vagueness_2017}, we can then define the following values.
Probability that $t_a$ is actual if $t_o$ is observed by an agent of type $\alpha$:
$$
P_{\bar{o}}^{\alpha}(t_{a}|t_{o})\propto P_{a}(t_{a})P_{o}^{\alpha}(t_{o}|t_{a})
$$
Probability that a random sender of type $\alpha$ produces $m$ when the actual state is $t_a$:
$$
P_{\sigma}^{\alpha}(m|t_{a})=\sum_{t_{o}}P_{o}^{\alpha}(t_{o}|t_{a})\sigma^{\alpha}(m|t_{o})
$$
Probability that the actual state is $t_a$ if a random sender of type $\alpha$ produced $m$:
$$
P_{\bar{\sigma}}^{\alpha}(t_{a}|m)\propto P_{a}(t_{a})P_{\sigma}^{\alpha}(m|t_{a})
$$
Probability that $t_r$ is realized by a random receiver of type $\alpha$ in response to message $m$:
$$
P_{\rho}^{\alpha}(t_{r}|m)=\sum_{t_{i}}P_{r}^{\alpha}(t_{r}|t_{i})\rho^{\alpha}(t_{i}|m)
$$
These formulations are merely parameterized versions of the single-population model.
They encapsulate calculations that we can use to compute expected utilities and update probabilities for each type.
The latter, however, depend on the types of interaction that we imagine occurring between populations.
In the following sections, we consider two different possibilities.

\subsection{Tight population interaction}
\label{sec:tight-interaction-model}
In a multi-population model with tight interaction between populations, each agent plays with, observes, and potentially imitates any other agent, regardless of their type.
This has an impact on the expected utilities of sender and receiver strategies of each type, and on the update probabilities for those strategies.
Let's start with the expected utilities.
For a sender of type $\alpha$, the expected utility of its strategy $\sigma^\alpha$ against all other receiver strategies $\rho^\star$, is given by:
$$
\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\star})=\sum_{t_{a}}P_{\bar{o}}^{\alpha}(t_{a}|t_{o})\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})\sum_{t_{r}}P_{\rho}^{\alpha^{\prime}}(t_{r}|m)U(t_{a},t_{r})
$$
and, for a receiver of type $\alpha$, the expected utility of its strategy $\rho^\alpha$ against all other senderstrategies $\sigma^\star$, is given by:
$$
\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\star})=\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})\sum_{t_{a}}P_{\bar{\sigma}}^{\alpha^{\prime}}(t_{a}|m)\sum_{t_{r}}P_{r}^{\alpha}(t_{r}|t_{i})U(t_{a},t_{r})
$$
Expected utilities thus take into account the existence of strategies of other types, and weigh the relevance of each type $\alpha^\prime$ according to its relative proportion $P(\alpha^\prime)$.

Another important value to calculate has to do with the types which agents observe and imitate.
In a model with tight interaction, we imagine this occurring across populations.
Therefore, we can define the probability that a sender of type $\alpha$ observes a randomly sampled agent play message $m$ for observed state $t_o$ as:
$$
P_{o}^{\alpha}(m|t_{o})=\sum_{t_{a}}P_{\bar{o}}^{\alpha}(t_{a}|t_{o})\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})P_{\sigma}^{\alpha^{\prime}}(m|t_{a})
$$
and the probability that a receiver of type $\alpha$ observes a randomly sampled agent choose interpretation $t_o$ given message $m$ as:
$$
P_{o}^{\alpha}(t_{o}|m)=\sum_{t_{r}}P_{o}^{\alpha}(t_{o}|t_{r})\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})P_{\rho}^{\alpha^{\prime}}(t_{r}|m)
$$
Again, these calculation incorporate the probabilities that the imitating agent might observe the behavior of an agent of another type $\alpha^\prime$, weighed by its relative proportion.
%
Finally, the update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ can be given by:
$$
\sigma_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)\text{EU}_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\star})
$$
and similarly for a receiver strategy of type $\alpha$ by:
$$
\rho_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)\text{EU}_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\star})
$$

@@

We model this process by changing the proportion of each type $P(\alpha)$ according to the replicator dynamic.
A population of type $\alpha$ consists of agents employing both a sender and a receiver strategy, thus the overall fitness of the population must include the expected utilities of both.
We could imagine this process happening at a different speed than the cultural process, in which case we could have different time scales.
For the sake of simplicity, we choose to have them both happen at each time step, but calculate the changes occurring at this level of selection after the calculation for the other level.
We define the proportion of type $\alpha$ at time step $i+1$ as:
$$
P_{i+1}(\alpha) \propto P_i(\alpha)(\text{EU}_{\sigma_{i+1}}^{\alpha}(m,t,\rho_{i}^{\star}) + \text{EU}_{\rho_{i+1}}^{\alpha}(t,m,\sigma_{i}^{\star}))
$$
In order to additionally account for the fact that strategies are not passed on to new generations, we mix the evolved strategies of a certain type $\sigma^\alpha$ and $\rho^\alpha$ with new random strategies $\tilde{\sigma}^\alpha$ and $\tilde{\rho}^\alpha$ (generated at each time step).
The idea is that, at each time step, a certain percentage of each population will consist of ``newborn'' agents, \emph{i.e.}~agents that haven't yet had time to evolve their strategies.
We define a parameter $\gamma$ that quantifies this percentage, or as we can also call it the birth rate, which we consider to be the same for every population.
This mixing can be described, for a certain time step $i+1$, in the following formulas:
$$
\sigma_{i+1}^{\alpha}(m|t)=(1 - \gamma)\sigma_{i+1}^{\alpha}(m|t) + \gamma \tilde{\sigma}_{i+1}^{\alpha}
$$
$$
\rho_{i+1}^{\alpha}(t|m)=(1 - \gamma)\rho_{i+1}^{\alpha}(t|m) + \gamma \tilde{\rho}_{i+1}^{\alpha}
$$

@@

The first thing to observe from the simulation results is that the population with no imprecision ($\alpha = 0$) dominated the other populations in every run.
%\begin{sidewaysfigure}
\begin{figure}
  \centering
  \begin{subfigure}[]{\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-strong.png}
    \caption{Two populations scenario.}
    \label{fig:proportion-cases-two-tight-interaction}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-strong.png}
    \caption{Three populations scenario.}
    \label{fig:proportion-cases-three-tight-interaction}
  \end{subfigure}
  \caption{Evolution of population proportions through time for each simulation trial of the tight interaction model.}
  \label{fig:proportion-cases-tight-interaction}
\end{figure}
%\end{sidewaysfigure}
In Figure~\ref{fig:proportion-cases-tight-interaction} we plot the evolution of the proportion of each population in the two-population and three-populations scenario for all trials.
As the plot shows, for every trial, the population with no imprecision steadily increased its proportion against the others.
In the three population scenario, the proportion of type $\alpha = 0.05$ sees a slight increase in the beginning of the simulation, but inexorably starts a downward trend.
These observations go against the expectation of~\textcite{franke_vagueness_2017} that faster convergence to a convex strategy by populations with a certain level of imprecision could give them a temporary advantage to take over and eliminate other types.
The reason for this is interesting in itself.
What happens is that, because of the tight interaction between populations, the strategies of each type evolve in close tandem with each other.
One of the consequences of this is that the population with no imprecision reaches convexity faster than it would on its own because of the interaction with the populations with imprecision.
We can see this effect by looking into the percentage of trials with convex sender strategies at a given iteration, for each scenario, and comparing the three scenarios: population with no imprecision evolving alone, two populations ($\alpha \in \{0, 0.05\}$), and three populations ($\alpha \in \{0, 0.05, 0.1\}$).
\begin{figure}
  \centering
  \begin{subfigure}[]{0.32\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/convex-percentage-all-strong.png}
    \caption{Percentage of convex sender strategies for $\alpha = 0$.}
    \label{fig:convexity-tight-interaction}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.32\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/entropy-sender-all-strong.png}
    \caption{Mean entropy of sender strategies for $\alpha = 0$.}
    \label{fig:entropy-sender-tight-interaction}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.32\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/entropy-receiver-all-strong.png}
    \caption{Mean entropy of receiver strategies for $\alpha = 0$.}
    \label{fig:entropy-receiver-tight-interaction}
  \end{subfigure}
  \caption{Development of some metrics through time for the $\alpha = 0$ population in each of three scenarios: evolving alone (`0'), with an $\alpha = 0.05$ population (`0/0.05'), and with an additional $\alpha = 0.1$ population (`0/0.05/0.1').}
  \label{fig:metrics-tight-interaction}
\end{figure}
This is plotted in Figure~\ref{fig:convexity-tight-interaction}.
What we see is more trials reaching convexity earlier for the multi-population scenarios when compared with the single-poulation scenario.
This effect precludes the hypothesized temporary advantage of imprecision to even manifest itself, but it can be seen as a positive influence on the population with no imprecision.

@@

Regarding the imitation process, we can redefine the probability that a sender of type $\alpha$ observes a randomly sampled agent play message $m$ for observed state $t_o$ as:
$$
P_{o}^{\alpha}(m|t_{o})=\sum_{t_{a}}P_{\bar{o}}^{\alpha}(t_{a}|t_{o})P_{\sigma}^{\alpha}(m|t_{a})
$$
and the probability that a receiver of type $\alpha$ observes a randomly sampled agent choose interpretation $t_o$ given message $m$ as:
$$
P_{o}^{\alpha}(t_{o}|m)=\sum_{t_{r}}P_{o}^{\alpha}(t_{o}|t_{r})P_{\rho}^{\alpha}(t_{r}|m)
$$
Again, the main difference is that agents make observations within their own population, so to model a randomly sampled agent one needs only to take into account agents of the same type.
Population proportions again do not play a role in these calculations.
Based on these formulas, we can define the update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ as:
$$
\sigma_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)\text{EU}_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\alpha})
$$
and similarly for a receiver strategy of type $\alpha$ as:
$$
\rho_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)\text{EU}_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\alpha})
$$
Note that, because imitation and learning occur only within populations, these formulations are essentially the same as for the single-population model of~\textcite{franke_vagueness_2017}, only parameterized by type $\alpha$.

@@

We ran the same number of simulation trials under the same conditions as for the model with tight population interaction.
%\begin{sidewaysfigure}
\begin{figure}
  \centering
  \begin{subfigure}[]{\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-weakest.png}
    \caption{Two populations scenario.}
    \label{fig:proportion-cases-two-loose-interaction}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-weakest.png}
    \caption{Three populations scenario.}
    \label{fig:proportion-cases-three-loose-interaction}
  \end{subfigure}
  \caption{Evolution of population proportions through time for each simulation trial of the loose interaction model.}
  \label{fig:proportion-cases-loose-interaction}
\end{figure}
%\end{sidewaysfigure}
In Figure~\ref{fig:proportion-cases-loose-interaction} we plot the evolution of population proportions for all trials of the multi-population scenarios.
The first thing to observe is that proportions evolve faster than in the model with tight interaction.
Whereas in the latter no given population ever reached much more than 70\% proportion, in this model we see that many trials resulted in one population fully dominating the others.
In the two-population scenario, some population reached 99\% proportion in 24 out of 25 trials.
In the three-population scenario, this happened in 18 out of 25 trials.
%Given that strategies no longer develop in close tandem with each other, differences in expected utility can become more salient resulting in some type reaching a more clear advantage than the other.
More interestingly for our investigation, some trials actually resulted in a population with some degree of imprecision dominating the others.
For the two-population scenario, $\alpha = 0.05$ fully reached 100\% proportion in 8 trials (79, 86, 88, 89, 91, 93, 96, 98), a point from which it is technically impossible for the other population to recover.
In most cases the dominating population gains its ground from the start, but in 5 cases we see a temporary advantage of $\alpha = 0.05$ that is then lost to the other population.
In one interesting trial (85), $\alpha = 0.05$ reached 99.87\% only to then steadly lose ground to $\alpha = 0$.
For the three population scenario, $\alpha = 0$ was reduced to 0\% proportion in 4 trials (152, 157, 159, and 175).
In all of those cases, $\alpha = 0.05$ clearly has the upper hand over $\alpha = 0.1$, despite a temporary advantage of the latter in some trials.
%In 5 other trials we see a positive trend for $\alpha = 0.05$, but given the reversals that sometimes occur (\emph{e.g.}~trial 171), one cannot be sure whether full dominance would be achieved if the simulation was left to run for longer time.

>>>>

In the following, we lay down the formal details of our model.
Let's start by defining $A$ to be the set of imprecision values considered.
For each value $\alpha \in A$, the proportion of its population is given by $P(\alpha)$, such that $\sum_{\alpha \in A} P(\alpha) = 1$.
Each population has its own sender and receiver strategies, represented as $\sigma^\alpha$ and $\rho^\alpha$.
The probability that a given agent with imprecision $\alpha$ (or of type $\alpha$) observes $t_o$ when the actual state is $t_a$ is given by $P_{o}^{\alpha}(t_{o}|t_{a})$.
If the same agent intends to realize $t_i$, the probability that he actually realizes $t_r$ instead is given by $P_{r}^{\alpha}(t_{r}|t_{i})$.
Following~\textcite[26]{franke_vagueness_2017}, we can then define the following values.
Probability that $t_a$ is actual if $t_o$ is observed by an agent of type $\alpha$:
$$
P_{\bar{o}}^{\alpha}(t_{a}|t_{o})\propto P_{a}(t_{a})P_{o}^{\alpha}(t_{o}|t_{a})
$$
Probability that a random sender of type $\alpha$ produces $m$ when the actual state is $t_a$:
$$
P_{\sigma}^{\alpha}(m|t_{a})=\sum_{t_{o}}P_{o}^{\alpha}(t_{o}|t_{a})\sigma^{\alpha}(m|t_{o})
$$
Probability that the actual state is $t_a$ if a random sender of type $\alpha$ produced $m$:
$$
P_{\bar{\sigma}}^{\alpha}(t_{a}|m)\propto P_{a}(t_{a})P_{\sigma}^{\alpha}(m|t_{a})
$$
Probability that $t_r$ is realized by a random receiver of type $\alpha$ in response to message $m$:
$$
P_{\rho}^{\alpha}(t_{r}|m)=\sum_{t_{i}}P_{r}^{\alpha}(t_{r}|t_{i})\rho^{\alpha}(t_{i}|m)
$$
These formulations are merely parameterized versions of the single-population model.
They encapsulate calculations that we can use to compute expected utilities and strategy update steps for each type.
The latter, however, depend on the types of interaction that we imagine occurring between populations.
In the following sections, we consider two different possibilities.

\subsection{Tight population interaction}
\label{sec:tight-interaction-model}
In a multi-population model with tight interaction between populations, each agent plays with, observes, and potentially imitates any other agent, regardless of their type.
This has an impact on the expected utilities of sender and receiver strategies of each type, and on the update steps for those strategies.
Let's start with the expected utilities.
For a sender of type $\alpha$, the expected utility of its strategy $\sigma^\alpha$ against all other receiver strategies $\rho^\star$, is given by:
$$
\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\star})=\sum_{t_{a}}P_{\bar{o}}^{\alpha}(t_{a}|t_{o})\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})\sum_{t_{r}}P_{\rho}^{\alpha^{\prime}}(t_{r}|m)U(t_{a},t_{r})
$$
and, for a receiver of type $\alpha$, the expected utility of its strategy $\rho^\alpha$ against all other senderstrategies $\sigma^\star$, is given by:
$$
\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\star})=\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})\sum_{t_{a}}P_{\bar{\sigma}}^{\alpha^{\prime}}(t_{a}|m)\sum_{t_{r}}P_{r}^{\alpha}(t_{r}|t_{i})U(t_{a},t_{r})
$$
Expected utilities thus take into account the existence of strategies of other types, and weigh the relevance of each type $\alpha^\prime$ according to its relative proportion $P(\alpha^\prime)$.

Another important value to calculate has to do with the types which agents observe and imitate.
In a model with tight interaction, we imagine this occurring across populations.
Therefore, we can define the probability that a sender of type $\alpha$ observes a randomly sampled agent play message $m$ for observed state $t_o$ as:
$$
P_{o}^{\alpha}(m|t_{o})=\sum_{t_{a}}P_{\bar{o}}^{\alpha}(t_{a}|t_{o})\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})P_{\sigma}^{\alpha^{\prime}}(m|t_{a})
$$
and the probability that a receiver of type $\alpha$ observes a randomly sampled agent choose interpretation $t_o$ given message $m$ as:
$$
P_{o}^{\alpha}(t_{o}|m)=\sum_{t_{r}}P_{o}^{\alpha}(t_{o}|t_{r})\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})P_{\rho}^{\alpha^{\prime}}(t_{r}|m)
$$
Again, these calculation incorporate the probabilities that the imitating agent might observe the behavior of an agent of another type $\alpha^\prime$, weighed by its relative proportion.
%
Finally, the update step for a sender strategy of type $\alpha$ at time instant $i+1$ is given by:
$$
\check{\sigma}_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)\text{EU}_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\star})
$$
and similarly for a receiver strategy of type $\alpha$ by:
$$
\check{\rho}_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)\text{EU}_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\star})
$$
We here use $\check{\sigma}$ and $\check{\rho}$ since there is still an additional adjustment to these values to be calculated before we get the final strategies $\sigma$ and $\rho$.

@@

We model this process by changing the proportion of each type $P(\alpha)$ according to the replicator dynamic.
A population of type $\alpha$ consists of agents employing both a sender and a receiver strategy, thus the overall fitness of the population must include the expected utilities of both.
We could imagine this process happening at a different speed than the cultural process, in which case we could have different time scales.
For the sake of simplicity, we choose to have them both happen at each time step, but calculate the changes occurring at this level of selection after the calculation for the other level.
We define the proportion of type $\alpha$ at time step $i+1$ as:
$$
P_{i+1}(\alpha) \propto P_i(\alpha)(\text{EU}_{\sigma_{i+1}}^{\alpha}(m,t,\rho_{i}^{\star}) + \text{EU}_{\rho_{i+1}}^{\alpha}(t,m,\sigma_{i}^{\star}))
$$
In order to additionally account for the fact that strategies are not passed on to new generations, we mix the evolved strategies of a certain type $\sigma^\alpha$ and $\rho^\alpha$ with new random strategies $\tilde{\sigma}^\alpha$ and $\tilde{\rho}^\alpha$ (generated at each time step).
The idea is that, at each time step, a certain percentage of each population will consist of ``newborn'' agents, \emph{i.e.}~agents that haven't yet had time to evolve their strategies.
We define a parameter $\gamma$ that quantifies this percentage, or as we can also call it the birth rate, which we consider to be the same for every population.
This mixing finally defines the strategies for time step $i+1$ and can be described in the following formulas:
$$
\sigma_{i+1}^{\alpha}(m|t)=(1 - \gamma)\check{\sigma}_{i+1}^{\alpha}(m|t) + \gamma \tilde{\sigma}_{i+1}^{\alpha}(m|t)
$$
$$
\rho_{i+1}^{\alpha}(t|m)=(1 - \gamma)\check{\rho}_{i+1}^{\alpha}(t|m) + \gamma \tilde{\rho}_{i+1}^{\alpha}(t|m)
$$

@@

The first thing to observe from the simulation results is that the population with no imprecision ($\alpha = 0$) dominated the other populations in every run.
%\begin{sidewaysfigure}
\begin{figure}
  \centering
  \begin{subfigure}[]{.9\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-strong-square.png}
    \caption{Two populations scenario.}
    \label{fig:proportion-cases-two-tight-interaction}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{.9\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-strong-square.png}
    \caption{Three populations scenario.}
    \label{fig:proportion-cases-three-tight-interaction}
  \end{subfigure}
  \caption{Evolution of population proportions through time for each simulation trial of the tight interaction model. Numbers on top of each plot identify each trial.}
  \label{fig:proportion-cases-tight-interaction}
\end{figure}
%\end{sidewaysfigure}
In Figure~\ref{fig:proportion-cases-tight-interaction} we plot the evolution of the proportion of each population in the two-population and three-populations scenario for all trials.
As the plot shows, for every trial, the population with no imprecision steadily increased its proportion against the others.
In the three population scenario, the proportion of type $\alpha = 0.05$ sees a slight increase in the beginning of the simulation, but inexorably starts a downward trend.
These observations go against the expectation of~\textcite{franke_vagueness_2017} that faster convergence to a convex strategy by populations with a certain level of imprecision could give them a temporary advantage to take over and eliminate other types.
The reason for this is interesting in itself.
What happens is that, because of the tight interaction between populations, the strategies of each type evolve in close tandem with each other.
One of the consequences of this is that the population with no imprecision reaches convexity faster than it would on its own because of the interaction with the populations with imprecision.
We can see this effect by looking into the percentage of trials with convex sender strategies at a given iteration, for each scenario, and comparing the three scenarios: population with no imprecision evolving alone, two populations ($\alpha \in \{0, 0.05\}$), and three populations ($\alpha \in \{0, 0.05, 0.1\}$).
\begin{figure}
  \centering
  \begin{subfigure}[]{0.32\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/convex-percentage-all-strong.png}
    \caption{Percentage of convex sender strategies for $\alpha = 0$.}
    \label{fig:convexity-tight-interaction}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.32\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/entropy-sender-all-strong.png}
    \caption{Mean entropy of sender strategies for $\alpha = 0$.}
    \label{fig:entropy-sender-tight-interaction}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.32\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/entropy-receiver-all-strong.png}
    \caption{Mean entropy of receiver strategies for $\alpha = 0$.}
    \label{fig:entropy-receiver-tight-interaction}
  \end{subfigure}
  \caption{Development of some metrics through time for the $\alpha = 0$ population in each of three scenarios: evolving alone (`0'), with an $\alpha = 0.05$ population (`0/0.05'), and with an additional $\alpha = 0.1$ population (`0/0.05/0.1').}
  \label{fig:metrics-tight-interaction}
\end{figure}
This is plotted in Figure~\ref{fig:convexity-tight-interaction}.
What we see is more trials reaching convexity earlier for the multi-population scenarios when compared with the single-poulation scenario.
This effect precludes the hypothesized temporary advantage of imprecision to even manifest itself, but it can be seen as a positive influence on the population with no imprecision.

@@

Regarding the imitation process, we can redefine the probability that a sender of type $\alpha$ observes a randomly sampled agent play message $m$ for observed state $t_o$ as:
$$
P_{o}^{\alpha}(m|t_{o})=\sum_{t_{a}}P_{\bar{o}}^{\alpha}(t_{a}|t_{o})P_{\sigma}^{\alpha}(m|t_{a})
$$
and the probability that a receiver of type $\alpha$ observes a randomly sampled agent choose interpretation $t_o$ given message $m$ as:
$$
P_{o}^{\alpha}(t_{o}|m)=\sum_{t_{r}}P_{o}^{\alpha}(t_{o}|t_{r})P_{\rho}^{\alpha}(t_{r}|m)
$$
Again, the main difference is that agents make observations within their own population, so to model a randomly sampled agent one needs only to take into account agents of the same type.
Population proportions again do not play a role in these calculations.
Based on these formulas, we can define the update step for a sender strategy of type $\alpha$ at time instant $i+1$ as:
$$
\check{\sigma}_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)\text{EU}_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\alpha})
$$
and similarly for a receiver strategy of type $\alpha$ as:
$$
\check{\rho}_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)\text{EU}_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\alpha})
$$
Note that, because imitation and learning occur only within populations, these formulations are essentially the same as for the single-population model of~\textcite{franke_vagueness_2017}, only parameterized by type $\alpha$.

@@

We ran the same number of simulation trials under the same conditions as for the model with tight population interaction.
%\begin{sidewaysfigure}
\begin{figure}
  \centering
  \begin{subfigure}[]{.9\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-weakest-square.png}
    \caption{Two populations scenario.}
    \label{fig:proportion-cases-two-loose-interaction}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{.9\textwidth}
    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-weakest-square.png}
    \caption{Three populations scenario.}
    \label{fig:proportion-cases-three-loose-interaction}
  \end{subfigure}
  \caption{Evolution of population proportions through time for each simulation trial of the loose interaction model. Numbers on top of each plot identify each trial.}
  \label{fig:proportion-cases-loose-interaction}
\end{figure}
%\end{sidewaysfigure}
In Figure~\ref{fig:proportion-cases-loose-interaction} we plot the evolution of population proportions for all trials of the multi-population scenarios.
The first thing to observe is that proportions evolve faster than in the model with tight interaction.
Whereas in the latter no given population ever reached much more than 70\% proportion, in this model we see that many trials resulted in one population fully dominating the others.
In the two-population scenario, some population reached 99\% proportion in 24 out of 25 trials.
In the three-population scenario, this happened in 18 out of 25 trials.
%Given that strategies no longer develop in close tandem with each other, differences in expected utility can become more salient resulting in some type reaching a more clear advantage than the other.
More interestingly for our investigation, some trials actually resulted in a population with some degree of imprecision dominating the others.
For the two-population scenario, $\alpha = 0.05$ fully reached 100\% proportion in 8 trials (79, 86, 88, 89, 91, 93, 96, 98), a point from which it is technically impossible for the other population to recover.
In most cases the dominating population gains its ground from the start, but in 5 cases we see a temporary advantage of $\alpha = 0.05$ that is then lost to the other population.
In one interesting trial (85), $\alpha = 0.05$ reached 99.87\% only to then steadly lose ground to $\alpha = 0$.
For the three population scenario, $\alpha = 0$ was reduced to 0\% proportion in 4 trials (152, 157, 159, and 175).
In all of those cases, $\alpha = 0.05$ clearly has the upper hand over $\alpha = 0.1$, despite a temporary advantage of the latter in some trials.
%In 5 other trials we see a positive trend for $\alpha = 0.05$, but given the reversals that sometimes occur (\emph{e.g.}~trial 171), one cannot be sure whether full dominance would be achieved if the simulation was left to run for longer time.
